Algorithmic Approaches to Match Degraded Land Impressions

### 2017-10-06 degraded-paper.Rmd
Small changes

In this paper, we take steps to address these and other concerns. Specifically, we begin by reviewing features from the literature, computed on pairs of bullet lands, and presenting some of our own features. We propose an approach to  standardize the featues, to account for the fact that only a portion of the land impression may be recovered from the crime scene. With the standardized features, we tackle two issues that were not addressed in  [@2016arXiv160105788H]. The first is the effect of the microscope operator  on the resulting images and consequent algorithm performance. The second issue has to do with the robustness of the land matching algorithm in [@2016arXiv160105788H] relative to the degree of degradation of the questioned land impression. Finally, we describe some of the initial steps toward generalizing a matching algorithm based on land-to-land comparisons, to one based on bullet-to-bullet comparisons, as would be of interest in a real world application of these ideas.

@@

```{r, echo=FALSE, results='asis'}
study_result <- do.call(rbind, result_study)
study_result <- cbind(id = rep(1:10, each = nrow(study_result) / 10), study_result)
study_result_wide <- study_result %>%
    spread(key = Actual, value = Count) %>%
    group_by(Prediction) %>%
    summarise_each(funs(mean)) %>%
    select(-id) %>%
    as.data.frame
study_result_wide_col <- study_result_wide
study_result_wide_col[,-1] <- apply(study_result_wide_col[,-1], 2, function(x) paste0(round(100 * x / sum(x), digits = 2), "%"))
names(study_result_wide_col)[1] <- "Prediction \\ Actual"

@@

Though visual inspection clearly shows differences, we can more formally assess the differences between distributions with a Kolmogrov-Smirnov test. Table \ref{tab:kstests} gives the results of pairwise tests, for each feature, between different set comparisons, and between known matches compared with known non-matches. Although most of the tests are significant, looking at the raw values of the D statistic suggest that the largest effect sizes do in fact occur in comparisons with two Hamby252 lands, as the visual inspection of the boxplots also suggested.

@@

print(xtable(final_result[,1:7], digits = c(0, 0, 0, 0, 0, 4, 0, 4), label = "tab:kstests", caption = "Results for the Kolmogrov-Smirnov distributional test."), comment = FALSE, include.rownames = FALSE, table.placement = 'H')
```

@@

print(xtable(tidy(mytest), digits = c(0, 4, 4, 0, 0), label = "tab:mykstest", caption = "Results for the Kolmogrov-Smirnov distributional test between values of the ideal cross section for Hamby252 compared with Hamby44."), comment = FALSE, include.rownames = FALSE)
```

@@

Figure \@ref(fig:senspe) gives the sensitivity (true positive rate) and specificity (true negative rate) of the random forest predictions for given levels of degradation. It can be seen that the sensitivity drops a bit until 50% of the land is available and then rises again. This occurs because the algorithm begins producing more positive predictions in general, likely as the result of the ccf being arbitrarily higher for known non-matches due to the small signature. On the other hand, the specificity drops dramatically for left, middle, and right fixed degraded lands when less than 50%. of the land impression is available for examination.  For a more in-depth exploration of the matching probabilities, Figure \@ref(fig:deghist) provides histograms of the matching probability  by  degradation level and by known match versus known non-match categories. The matching probabilities suffer when compared with the probablities obtained from comparisons between full land impressions in all cases. The jump seems to be most noticeable beginning at about 25% degradation (75% land recovered), and the algorithm struggles beyond 50%.

@@

\begin{figure}[H]
\centering
\includegraphics[width=\linewidth]{images/br9-2-4-greyflip.png}
\caption{Land 4 of Bullet 2, from Barrel 9 of Hamby Set 252. It can be seen that this particular land exhibits some major tank rash on the right half.}
\label{fig:br924}
\end{figure}

Table \ref{tab:br924pred} shows the values of the features, after extracting only the first 50% of the Hamby Barrel 9 Bullet 2, 4th land (and hence, simulating a left-fixed 50% degraded scenario), compared with a feature comparison between both full lands (and hence, including the tank rash striae). The features are derived in a comparison with its known match, the complete Bullet 1 third land fired from Barrel 9. The features, including the ccf and the matches, are expressed enough to (barely) indicate a match in the case of the degraded bullet. Using the pre-trained random forest, the predicted matching probability is 52%. This is encouraging in that attempting to match the full bullet land, by comparison, yields a matching probability of 0.0067%. This is due to the relatively higher values of the ccf, cms, and matches for the degraded comparison, and suggests that the feature standardization is working as intended.

@@

where $M$ is the event that two bullets match, $NM$ is the event that two bullets do not match, $M1$, $M2$, ..., $M6$ are the probabilities of land one, land two, ..., land six matching, and $NM1$, $NM2$, ..., $NM6$ are the probabilities that land one, land two, ..., land six do not match. However, to compute this probability, we need to know the alignment of the two sets of lands. Fortunately, the consistent rotation of the bullet permits this. For instance, if we knew that land 1 of bullet 1 matches land 4 of bullet 2, then we immediately know that land 2 of bullet 1 matches to land 5 of bullet 2, land 3 of bullet 1 matches to land 6 of bullet 2, etc. Hence, we can take look across six diagonals of the $6 \otimes 6$ matrix containing match probabilities. Table \ref{tab:diag} gives an example of the matrix of matching probabilities between two sets of six lands from bullets that are known matches. The matching diagonal is clear based on the high probabilities (cell $(1, 3)$, cell $(2, 4)$, cell $(3, 5)$, etc.) although it can be seen that one of the six comparisons has a relatively lower matching probability. This procedure is based on the Sequence Average Maximum (SAM) by @sensorfar in their bullet matching software application `SensoMatch`. A similar approach using the cross correlation maximum was first proposed by @chu:2010. Compared to that approach, ours uses random forest based probabilities compared to correlation values allowing for elements of probability theory to help determine the resulting bullet match probability.

@@

We began exploring the robustness of the matching algorithm proposed in [@2016arXiv160105788H] to land degradation. As suspected, the algorithm performance declines as a function of the rate of degradation. However, there is a relatively clear threshold around about 50%; if 50% of the land or more is recovered, the algorithm still performs reasonably well.  When the proportion of the land that is recovered is below 50%, the accuracy with which we can compare land impressions is low.

>>>>

In this paper, we take steps to address these and other concerns. Specifically, we begin by reviewing features from the literature, computed on pairs of bullet lands, and presenting some of our own features. We propose an approach to  standardize the featues, to account for the fact that only a portion of the land impression may be recovered from the crime scene. With the standardized features, we tackle two issues that were not addressed in  @2016arXiv160105788H. The first is the effect of the microscope operator  on the resulting images and consequent algorithm performance. The second issue has to do with the robustness of the land matching algorithm in @2016arXiv160105788H relative to the degree of degradation of the questioned land impression. Finally, we describe some of the initial steps toward generalizing a matching algorithm based on land-to-land comparisons, to one based on bullet-to-bullet comparisons, as would be of interest in a real world application of these ideas.

@@

```{r, echo=FALSE, results='asis'}
study_result <- do.call(rbind, result_study)
study_result <- cbind(id = rep(1:10, each = nrow(study_result) / 10), study_result)
study_result_wide <- study_result %>%
    spread(key = Actual, value = Count) %>%
    group_by(Prediction) %>%
    summarise_each(funs(mean)) %>%
    select(-id) %>%
    as.data.frame
study_result_wide_col <- study_result_wide
study_result_wide_col[,-1] <- apply(study_result_wide_col[,-1], 2, function(x) {
    mystr <- paste0(round(100 * x / sum(x), digits = 2), "%")
    mystr[stringr::str_length(mystr) %in% c(4, 5)] <- paste0("0", mystr[stringr::str_length(mystr) %in% c(4, 5)])

    return(mystr)
})
names(study_result_wide_col)[1] <- "Prediction \\ Actual"

@@

Though visual inspection clearly shows differences, we can more formally assess the differences between distributions with a Kolmogorov-Smirnov test. Table \ref{tab:kstests} gives the results of pairwise tests, for each feature, between different set comparisons, and between known matches compared with known non-matches. Although most of the tests are significant, looking at the raw values of the D statistic suggest that the largest effect sizes do in fact occur in comparisons with two Hamby252 lands, as the visual inspection of the boxplots also suggested.

@@

print(xtable(final_result[,1:7], digits = c(0, 0, 0, 0, 0, 4, 0, 4), label = "tab:kstests", caption = "Results for the Kolmogorov-Smirnov distributional test."), comment = FALSE, include.rownames = FALSE, table.placement = 'H')
```

@@

print(xtable(tidy(mytest), digits = c(0, 4, 4, 0, 0), label = "tab:mykstest", caption = "Results for the Kolmogorov-Smirnov distributional test between values of the ideal cross section for Hamby252 compared with Hamby44."), comment = FALSE, include.rownames = FALSE)
```

@@

Figure \@ref(fig:senspe) gives the sensitivity (true positive rate) and specificity (true negative rate) of the random forest predictions for given levels of degradation. It can be seen that the sensitivity drops a bit until 50% of the land is available and then rises again. This occurs because the algorithm begins producing more positive predictions in general, likely as the result of the ccf being arbitrarily higher for known non-matches due to the small signature. On the other hand, the specificity drops dramatically for left, middle, and right fixed degraded lands when less than 50% of the land impression is available for examination.  For a more in-depth exploration of the matching probabilities, Figure \@ref(fig:deghist) provides histograms of the matching probability  by  degradation level and by known match versus known non-match categories. The matching probabilities suffer when compared with the probablities obtained from comparisons between full land impressions in all cases. The jump seems to be most noticeable beginning at about 25% degradation (75% land recovered), and the algorithm struggles beyond 50%.

@@

\begin{figure}[H]
\centering
\includegraphics[width=\linewidth]{images/br9-2-4-grey.png}
\caption{Land 4 of Bullet 2, from Barrel 9 of Hamby Set 252. It can be seen that this particular land exhibits some major tank rash on the left half.}
\label{fig:br924}
\end{figure}

Table \ref{tab:br924pred} shows the values of the features, after extracting only the last 50% of the Hamby Barrel 9 Bullet 2, 4th land (and hence, simulating a right-fixed 50% degraded scenario), compared with a feature comparison between both full lands (and hence, including the tank rash striae). The features are derived in a comparison with its known match, the complete Bullet 1 third land fired from Barrel 9. The features, including the ccf and the matches, are expressed enough to (barely) indicate a match in the case of the degraded bullet. Using the pre-trained random forest, the predicted matching probability is 52%. This is encouraging in that attempting to match the full bullet land, by comparison, yields a matching probability of 0.0067%. This is due to the relatively higher values of the ccf, cms, and matches for the degraded comparison, and suggests that the feature standardization is working as intended.

@@

where $M$ is the event that two bullets match, $NM$ is the event that two bullets do not match, $M1$, $M2$, ..., $M6$ are the probabilities of land one, land two, ..., land six matching, and $NM1$, $NM2$, ..., $NM6$ are the probabilities that land one, land two, ..., land six do not match. However, to compute this probability, we need to know the alignment of the two sets of lands. Fortunately, the consistent rotation of the bullet permits this. For instance, if we knew that land 1 of bullet 1 matches land 4 of bullet 2, then we immediately know that land 2 of bullet 1 matches to land 5 of bullet 2, land 3 of bullet 1 matches to land 6 of bullet 2, etc. Hence, we can take look across six diagonals of the $6 \otimes 6$ matrix containing match probabilities. Table \ref{tab:diag} gives an example of the matrix of matching probabilities between two sets of six lands from bullets that are known matches. The matching diagonal is clear based on the high probabilities (cell $(1, 3)$, cell $(2, 4)$, cell $(3, 5)$, etc.) although it can be seen that one of the six comparisons has a relatively lower matching probability. This procedure is based on the Sequence Average Maximum (SAM) by @sensofar in their bullet matching software application `SensoMatch`. A similar approach using the cross correlation maximum was first proposed by @chu:2010. Compared to that approach, ours uses random forest based probabilities compared to correlation values allowing for elements of probability theory to help determine the resulting bullet match probability.

@@

We began exploring the robustness of the matching algorithm proposed in @2016arXiv160105788H to land degradation. As suspected, the algorithm performance declines as a function of the rate of degradation. However, there is a relatively clear threshold around about 50%; if 50% of the land or more is recovered, the algorithm still performs reasonably well.  When the proportion of the land that is recovered is below 50%, the accuracy with which we can compare land impressions is low.

### 2017-10-06 degraded-paper.tex
Small changes

Intense scrutiny has been focused on the process of bullet matching in
recent years (e.g., Giannelli 2011). Bullet matching, the process of
determining whether two bullets could have been fired from the same gun
barrel, has traditionally been performed without meaningful
determination of error rates or statistical assessments of uncertainty
(National Research Council 2009). There have been some attempts towards
developing mathematical and statistical approaches to bullet matching.
One such attempt was the definition of CMS, the Consecutively Matching
Striae (Biasotti 1959), with a cutoff of six to separate matches from
non-matches. Still, rigorous assessments of the applicability of such
cutoffs have not to this point been described (Advisors on Science and
Technology 2016).

@@

Real world applications of bullet matching often involve the recovery of
fragments of bullets from the crime scene (Council and others 2004).
Traditional features used in forensic examination work well for a full
land, but there has been less investigation into their performance in
the case of a fragmented land. For example, the CMS is naturally limited
by the portion of the land that can be recovered and varies across
manufacturers (W. Chu et al. 2011).

In this paper, we take steps to address these and other concerns.
Specifically, we begin by reviewing features from the literature,
computed on pairs of bullet lands, and presenting some of our own
features. We propose an approach to standardize the featues, to account
for the fact that only a portion of the land impression may be recovered
from the crime scene. With the standardized features, we tackle two
issues that were not addressed in (Hare, Hofmann, and Carriquiry 2016).
The first is the effect of the microscope operator on the resulting
images and consequent algorithm performance. The second issue has to do
with the robustness of the land matching algorithm in (Hare, Hofmann,
and Carriquiry 2016) relative to the degree of degradation of the
questioned land impression. Finally, we describe some of the initial
steps toward generalizing a matching algorithm based on land-to-land
comparisons, to one based on bullet-to-bullet comparisons, as would be
of interest in a real world application of these ideas.

@@

\begin{table}[H]
\centering
\begin{tabular}{llll}
  \hline
Prediction $\backslash$ Actual & Hamby252\_Hamby252 & Hamby252\_Hamby44 & Hamby44\_Hamby44 \\ 
  \hline
Hamby252\_Hamby252 & 9.6\% & 8.32\% & 11.28\% \\ 
  Hamby252\_Hamby44 & 81.47\% & 81.06\% & 78.68\% \\ 
  Hamby44\_Hamby44 & 8.93\% & 10.63\% & 10.04\% \\ 
   \hline
\end{tabular}
\caption{Confusion Matrix (Column Proportions) for the random forest with study as the response. It can be seen that the random forest performs poorly, as hoped, indicating that a simple model to predict the study using the features available is not enough to detect the operator effects.} 
\label{tab:studypred}
\end{table}

@@

Though visual inspection clearly shows differences, we can more formally
assess the differences between distributions with a Kolmogrov-Smirnov
test. Table \ref{tab:kstests} gives the results of pairwise tests, for
each feature, between different set comparisons, and between known
matches compared with known non-matches. Although most of the tests are
significant, looking at the raw values of the D statistic suggest that
the largest effect sizes do in fact occur in comparisons with two
Hamby252 lands, as the visual inspection of the boxplots also suggested.

\begin{table}[H]
\centering
\begin{tabular}{llllrlr}
  \hline
set1 & set2 & feature & matchtest & matchd & nonmatchtest & nonmatchd \\ 
  \hline
H252\_H252 & H252\_H44 & ccf & $<$ 0.0001 & 0.2723 & 1e-04 & 0.0189 \\ 
  H252\_H252 & H252\_H44 & cms & $<$ 0.0001 & 0.1751 & $<$ 0.0001 & 0.0245 \\ 
  H252\_H252 & H252\_H44 & D & $<$ 0.0001 & 0.2567 & $<$ 0.0001 & 0.1049 \\ 
  H252\_H252 & H252\_H44 & matches & $<$ 0.0001 & 0.1933 & $<$ 0.0001 & 0.0327 \\ 
  H252\_H252 & H252\_H44 & mismatches & $<$ 0.0001 & 0.2015 & 0.3537 & 0.0079 \\ 
  H252\_H252 & H252\_H44 & overlap & 0.0492 & 0.0984 & $<$ 0.0001 & 0.0276 \\ 
  H252\_H252 & H252\_H44 & rough\_cor & $<$ 0.0001 & 0.2647 & $<$ 0.0001 & 0.0970 \\ 
  H252\_H252 & H252\_H44 & sum\_peaks & 8e-04 & 0.1426 & 0.0015 & 0.0162 \\ 
  H252\_H252 & H44\_H44 & ccf & $<$ 0.0001 & 0.2160 & $<$ 0.0001 & 0.0257 \\ 
  H252\_H252 & H44\_H44 & cms & $<$ 0.0001 & 0.2515 & $<$ 0.0001 & 0.0467 \\ 
  H252\_H252 & H44\_H44 & D & $<$ 0.0001 & 0.2342 & $<$ 0.0001 & 0.1946 \\ 
  H252\_H252 & H44\_H44 & matches & $<$ 0.0001 & 0.2770 & $<$ 0.0001 & 0.0713 \\ 
  H252\_H252 & H44\_H44 & mismatches & $<$ 0.0001 & 0.2505 & 0.0414 & 0.0138 \\ 
  H252\_H252 & H44\_H44 & overlap & 0.2432 & 0.0906 & $<$ 0.0001 & 0.0408 \\ 
  H252\_H252 & H44\_H44 & rough\_cor & $<$ 0.0001 & 0.2242 & $<$ 0.0001 & 0.1718 \\ 
  H252\_H252 & H44\_H44 & sum\_peaks & 1e-04 & 0.1926 & $<$ 0.0001 & 0.0289 \\ 
  H252\_H44 & H44\_H44 & ccf & 0.1149 & 0.0883 & $<$ 0.0001 & 0.0259 \\ 
  H252\_H44 & H44\_H44 & cms & 0.111 & 0.0888 & $<$ 0.0001 & 0.0262 \\ 
  H252\_H44 & H44\_H44 & D & 0.2923 & 0.0724 & $<$ 0.0001 & 0.0906 \\ 
  H252\_H44 & H44\_H44 & matches & 0.0603 & 0.0977 & $<$ 0.0001 & 0.0423 \\ 
  H252\_H44 & H44\_H44 & mismatches & 0.3301 & 0.0700 & 0.1633 & 0.0096 \\ 
  H252\_H44 & H44\_H44 & overlap & 0.8231 & 0.0465 & 1e-04 & 0.0190 \\ 
  H252\_H44 & H44\_H44 & rough\_cor & 0.2671 & 0.0741 & $<$ 0.0001 & 0.0769 \\ 
  H252\_H44 & H44\_H44 & sum\_peaks & 0.047 & 0.1011 & 0.006 & 0.0147 \\ 
   \hline
\end{tabular}
\caption{Results for the Kolmogrov-Smirnov distributional test.} 
\label{tab:kstests}
\end{table}

@@

Figure \ref{fig:senspe} gives the sensitivity (true positive rate) and
specificity (true negative rate) of the random forest predictions for
given levels of degradation. It can be seen that the sensitivity drops a
bit until 50\% of the land is available and then rises again. This
occurs because the algorithm begins producing more positive predictions
in general, likely as the result of the ccf being arbitrarily higher for
known non-matches due to the small signature. On the other hand, the
specificity drops dramatically for left, middle, and right fixed
degraded lands when less than 50\%. of the land impression is available
for examination. For a more in-depth exploration of the matching
probabilities, Figure \ref{fig:deghist} provides histograms of the
matching probability by degradation level and by known match versus
known non-match categories. The matching probabilities suffer when
compared with the probablities obtained from comparisons between full
land impressions in all cases. The jump seems to be most noticeable
beginning at about 25\% degradation (75\% land recovered), and the
algorithm struggles beyond 50\%.

@@

\begin{figure}[H]
\centering
\includegraphics[width=\linewidth]{images/br9-2-4-greyflip.png}
\caption{Land 4 of Bullet 2, from Barrel 9 of Hamby Set 252. It can be seen that this particular land exhibits some major tank rash on the right half.}
\label{fig:br924}
\end{figure}

Table \ref{tab:br924pred} shows the values of the features, after
extracting only the first 50\% of the Hamby Barrel 9 Bullet 2, 4th land
(and hence, simulating a left-fixed 50\% degraded scenario), compared
with a feature comparison between both full lands (and hence, including
the tank rash striae). The features are derived in a comparison with its
known match, the complete Bullet 1 third land fired from Barrel 9. The
features, including the ccf and the matches, are expressed enough to
(barely) indicate a match in the case of the degraded bullet. Using the
pre-trained random forest, the predicted matching probability is 52\%.
This is encouraging in that attempting to match the full bullet land, by
comparison, yields a matching probability of 0.0067\%. This is due to
the relatively higher values of the ccf, cms, and matches for the
degraded comparison, and suggests that the feature standardization is
working as intended.

@@

We began exploring the robustness of the matching algorithm proposed in
(Hare, Hofmann, and Carriquiry 2016) to land degradation. As suspected,
the algorithm performance declines as a function of the rate of
degradation. However, there is a relatively clear threshold around about
50\%; if 50\% of the land or more is recovered, the algorithm still
performs reasonably well. When the proportion of the land that is
recovered is below 50\%, the accuracy with which we can compare land
impressions is low.

@@

\hypertarget{refs}{}
\hypertarget{ref-pcast2016}{}
Advisors on Science, President's Council of, and Technology. 2016.
``Report on Forensic Science in Criminal Courts: Ensuring Scientific
Validity of Feature-Comparison Methods.''
\url{https://www.whitehouse.gov/sites/default/files/microsites/ostp/PCAST/pcast_forensic_science_report_final.pdf}.

\hypertarget{ref-biasotti:1959}{}
Biasotti, Alfred A. 1959. ``A Statistical Study of the Individual
Characteristics of Fired Bullets.'' \emph{Journal of Forensic Sciences}
4 (1): 34--50.

@@

\hypertarget{ref-national2004forensic}{}
Council, National Research, and others. 2004. \emph{Forensic Analysis:
Weighing Bullet Lead Evidence}. National Academies Press.

@@

\hypertarget{ref-NAS:2009}{}
National Research Council. 2009. \emph{Strengthening Forensic Science in
the United States: A Path Forward}. Washington, DC: The National
Academies Press.
doi:\href{https://doi.org/10.17226/12589}{10.17226/12589}.

@@

@@

\hypertarget{ref-sensorfar}{}
Sensorfar. 2017. \emph{SensoMATCH Bullet Comparison Software}.

>>>>

Intense scrutiny has been focused on the process of bullet matching in
recent years (e.g., Giannelli 2011). Bullet matching, the process of
determining whether two bullets could have been fired from the same gun
barrel, has traditionally been performed without meaningful
determination of error rates or statistical assessments of uncertainty
(National Research Council 2009). There have been some attempts towards
developing mathematical and statistical approaches to bullet matching.
One such attempt was the definition of CMS, the Consecutively Matching
Striae (Biasotti 1959), with a cutoff of six to separate matches from
non-matches. Still, rigorous assessments of the applicability of such
cutoffs have not to this point been described (President's Council of
Advisors on Science and Technology 2016).

@@

Real world applications of bullet matching often involve the recovery of
fragments of bullets from the crime scene (National Research Council
2004). Traditional features used in forensic examination work well for a
full land, but there has been less investigation into their performance
in the case of a fragmented land. For example, the CMS is naturally
limited by the portion of the land that can be recovered and varies
across manufacturers (W. Chu et al. 2011).

In this paper, we take steps to address these and other concerns.
Specifically, we begin by reviewing features from the literature,
computed on pairs of bullet lands, and presenting some of our own
features. We propose an approach to standardize the featues, to account
for the fact that only a portion of the land impression may be recovered
from the crime scene. With the standardized features, we tackle two
issues that were not addressed in Hare, Hofmann, and Carriquiry (2016).
The first is the effect of the microscope operator on the resulting
images and consequent algorithm performance. The second issue has to do
with the robustness of the land matching algorithm in Hare, Hofmann, and
Carriquiry (2016) relative to the degree of degradation of the
questioned land impression. Finally, we describe some of the initial
steps toward generalizing a matching algorithm based on land-to-land
comparisons, to one based on bullet-to-bullet comparisons, as would be
of interest in a real world application of these ideas.

@@

\begin{table}[H]
\centering
\begin{tabular}{llll}
  \hline
Prediction $\backslash$ Actual & Hamby252\_Hamby252 & Hamby252\_Hamby44 & Hamby44\_Hamby44 \\ 
  \hline
Hamby252\_Hamby252 & 09.6\% & 08.32\% & 11.28\% \\ 
  Hamby252\_Hamby44 & 81.47\% & 81.06\% & 78.68\% \\ 
  Hamby44\_Hamby44 & 08.93\% & 10.63\% & 10.04\% \\ 
   \hline
\end{tabular}
\caption{Confusion Matrix (Column Proportions) for the random forest with study as the response. It can be seen that the random forest performs poorly, as hoped, indicating that a simple model to predict the study using the features available is not enough to detect the operator effects.} 
\label{tab:studypred}
\end{table}

@@

Though visual inspection clearly shows differences, we can more formally
assess the differences between distributions with a Kolmogorov-Smirnov
test. Table \ref{tab:kstests} gives the results of pairwise tests, for
each feature, between different set comparisons, and between known
matches compared with known non-matches. Although most of the tests are
significant, looking at the raw values of the D statistic suggest that
the largest effect sizes do in fact occur in comparisons with two
Hamby252 lands, as the visual inspection of the boxplots also suggested.

\begin{table}[H]
\centering
\begin{tabular}{llllrlr}
  \hline
set1 & set2 & feature & matchtest & matchd & nonmatchtest & nonmatchd \\ 
  \hline
H252\_H252 & H252\_H44 & ccf & $<$ 0.0001 & 0.2723 & 1e-04 & 0.0189 \\ 
  H252\_H252 & H252\_H44 & cms & $<$ 0.0001 & 0.1751 & $<$ 0.0001 & 0.0245 \\ 
  H252\_H252 & H252\_H44 & D & $<$ 0.0001 & 0.2567 & $<$ 0.0001 & 0.1049 \\ 
  H252\_H252 & H252\_H44 & matches & $<$ 0.0001 & 0.1933 & $<$ 0.0001 & 0.0327 \\ 
  H252\_H252 & H252\_H44 & mismatches & $<$ 0.0001 & 0.2015 & 0.3537 & 0.0079 \\ 
  H252\_H252 & H252\_H44 & overlap & 0.0492 & 0.0984 & $<$ 0.0001 & 0.0276 \\ 
  H252\_H252 & H252\_H44 & rough\_cor & $<$ 0.0001 & 0.2647 & $<$ 0.0001 & 0.0970 \\ 
  H252\_H252 & H252\_H44 & sum\_peaks & 8e-04 & 0.1426 & 0.0015 & 0.0162 \\ 
  H252\_H252 & H44\_H44 & ccf & $<$ 0.0001 & 0.2160 & $<$ 0.0001 & 0.0257 \\ 
  H252\_H252 & H44\_H44 & cms & $<$ 0.0001 & 0.2515 & $<$ 0.0001 & 0.0467 \\ 
  H252\_H252 & H44\_H44 & D & $<$ 0.0001 & 0.2342 & $<$ 0.0001 & 0.1946 \\ 
  H252\_H252 & H44\_H44 & matches & $<$ 0.0001 & 0.2770 & $<$ 0.0001 & 0.0713 \\ 
  H252\_H252 & H44\_H44 & mismatches & $<$ 0.0001 & 0.2505 & 0.0414 & 0.0138 \\ 
  H252\_H252 & H44\_H44 & overlap & 0.2432 & 0.0906 & $<$ 0.0001 & 0.0408 \\ 
  H252\_H252 & H44\_H44 & rough\_cor & $<$ 0.0001 & 0.2242 & $<$ 0.0001 & 0.1718 \\ 
  H252\_H252 & H44\_H44 & sum\_peaks & 1e-04 & 0.1926 & $<$ 0.0001 & 0.0289 \\ 
  H252\_H44 & H44\_H44 & ccf & 0.1149 & 0.0883 & $<$ 0.0001 & 0.0259 \\ 
  H252\_H44 & H44\_H44 & cms & 0.111 & 0.0888 & $<$ 0.0001 & 0.0262 \\ 
  H252\_H44 & H44\_H44 & D & 0.2923 & 0.0724 & $<$ 0.0001 & 0.0906 \\ 
  H252\_H44 & H44\_H44 & matches & 0.0603 & 0.0977 & $<$ 0.0001 & 0.0423 \\ 
  H252\_H44 & H44\_H44 & mismatches & 0.3301 & 0.0700 & 0.1633 & 0.0096 \\ 
  H252\_H44 & H44\_H44 & overlap & 0.8231 & 0.0465 & 1e-04 & 0.0190 \\ 
  H252\_H44 & H44\_H44 & rough\_cor & 0.2671 & 0.0741 & $<$ 0.0001 & 0.0769 \\ 
  H252\_H44 & H44\_H44 & sum\_peaks & 0.047 & 0.1011 & 0.006 & 0.0147 \\ 
   \hline
\end{tabular}
\caption{Results for the Kolmogorov-Smirnov distributional test.} 
\label{tab:kstests}
\end{table}

@@

Figure \ref{fig:senspe} gives the sensitivity (true positive rate) and
specificity (true negative rate) of the random forest predictions for
given levels of degradation. It can be seen that the sensitivity drops a
bit until 50\% of the land is available and then rises again. This
occurs because the algorithm begins producing more positive predictions
in general, likely as the result of the ccf being arbitrarily higher for
known non-matches due to the small signature. On the other hand, the
specificity drops dramatically for left, middle, and right fixed
degraded lands when less than 50\% of the land impression is available
for examination. For a more in-depth exploration of the matching
probabilities, Figure \ref{fig:deghist} provides histograms of the
matching probability by degradation level and by known match versus
known non-match categories. The matching probabilities suffer when
compared with the probablities obtained from comparisons between full
land impressions in all cases. The jump seems to be most noticeable
beginning at about 25\% degradation (75\% land recovered), and the
algorithm struggles beyond 50\%.

@@

\begin{figure}[H]
\centering
\includegraphics[width=\linewidth]{images/br9-2-4-grey.png}
\caption{Land 4 of Bullet 2, from Barrel 9 of Hamby Set 252. It can be seen that this particular land exhibits some major tank rash on the left half.}
\label{fig:br924}
\end{figure}

Table \ref{tab:br924pred} shows the values of the features, after
extracting only the last 50\% of the Hamby Barrel 9 Bullet 2, 4th land
(and hence, simulating a right-fixed 50\% degraded scenario), compared
with a feature comparison between both full lands (and hence, including
the tank rash striae). The features are derived in a comparison with its
known match, the complete Bullet 1 third land fired from Barrel 9. The
features, including the ccf and the matches, are expressed enough to
(barely) indicate a match in the case of the degraded bullet. Using the
pre-trained random forest, the predicted matching probability is 52\%.
This is encouraging in that attempting to match the full bullet land, by
comparison, yields a matching probability of 0.0067\%. This is due to
the relatively higher values of the ccf, cms, and matches for the
degraded comparison, and suggests that the feature standardization is
working as intended.

@@

We began exploring the robustness of the matching algorithm proposed in
Hare, Hofmann, and Carriquiry (2016) to land degradation. As suspected,
the algorithm performance declines as a function of the rate of
degradation. However, there is a relatively clear threshold around about
50\%; if 50\% of the land or more is recovered, the algorithm still
performs reasonably well. When the proportion of the land that is
recovered is below 50\%, the accuracy with which we can compare land
impressions is low.

@@

\hypertarget{refs}{}
\hypertarget{ref-biasotti:1959}{}
Biasotti, Alfred A. 1959. ``A Statistical Study of the Individual
Characteristics of Fired Bullets.'' \emph{Journal of Forensic Sciences}
4 (1): 34--50.

@@

@@

\hypertarget{ref-national2004forensic}{}
National Research Council. 2004. \emph{Forensic Analysis: Weighing
Bullet Lead Evidence}. National Academies Press.

\hypertarget{ref-NAS:2009}{}
---------. 2009. \emph{Strengthening Forensic Science in the United
States: A Path Forward}. Washington, DC: The National Academies Press.
doi:\href{https://doi.org/10.17226/12589}{10.17226/12589}.

@@

\hypertarget{ref-pcast2016}{}
President's Council of Advisors on Science and Technology. 2016.
``Report on Forensic Science in Criminal Courts: Ensuring Scientific
Validity of Feature-Comparison Methods.''
\url{https://www.whitehouse.gov/sites/default/files/microsites/ostp/PCAST/pcast_forensic_science_report_final.pdf}.

@@

\hypertarget{ref-sensofar}{}
Sensorfar. 2017. \emph{SensoMATCH Bullet Comparison Software}.
